Rice yield improvements through plant breeding are offset by inherent yield declines over time

### 2018-03-11 output/breed_paper_revised.Rnw
Final edits.

To test the hypothesis, the influence of breeding and yield decline over time
were quantified on yield over time using three nested Bayesian hierarchical
models. First, yield was modeled as a response to the number of years since
the cultivar was officially released and the year of release using a
mixed-effects linear regression. To account for similarities between
years and between cultivars, year and cultivar were included in the
model as random effects.

@@

Where $yield_i$ is the grain yield for cultivar $k$ in plot $i$ during
year $j$,
the Intercept is the average yield for the RES across cultivars and years, $Rate_{yield\
decline}$ is the rate of yield decline per
${years\ since\ release_{ki}}$ for cultivar $k$ in plot $i$
(mean-centered, i.e., negative for years prior to release, zero at the year of release),
$Rate_{yield\ improvement}$ is the rate of yield improvement per
year (${release\ year_{ik}}$), $\beta_{year_j}$ is the random effect of
year $j$, and $\beta_{cultivar_k}$ is the random effect of cultivar
$k$, and lastly $\beta_{cultivar_k:year_j}$ is the interaction term
for cultivar $k$ in year $j$. This
full model was modified to further explore yield dynamics over
time. For comparison to estimates where yield decline over time is not taken
into account, the model was used as described above except omitting
the $Rate_{yield\ decline}$ term (Model 2). Lastly, a model was fit
to estimate if there were differences between cultivars in the rate of
yield decline over time by adding allowing the rate of decline to vary
by cultivar (i.e., adding
a $Rate_{yield\ decline_k} * years\ since\ release_{ki}$ term to the
model for the rate of yield decline for cultivar $k$) (Model 3).

@@

There was general agreement between the estimates from the Bayesian
and classical statistical approaches for all models, with the
estimates being similar between similar models. The Bayesian estimates
are presented here due to the inherent flexibility of Bayesian
inference, though the classical estimates can be found in the
supplemental material. Comparing the cultivars' yield during each's
initial time (average 3 years) in the experimental trials showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M\=/207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1} (M\=/205)
(Fig \ref{fig:first_yrs}). The relationship between the year
a cultivar was released and yield over the 33-year study period was
estimated to be slightly negative (\Sexpr{release_yr_coef}
kg ha\tsups{-1} yr\tsups{-1}) and does not provide
evidence of genetic yield improvements during this
period. Individual exceptions to this were two cultivars (M\=/204 and
M\=/205) that had higher yields than the others.

@@

Based on Model 3, there was no evidence of consistent differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M\=/206 over M\=/202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M\=/202's yield over time (Figs \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M\=/202 and M\=/205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M\=/202 was
released 19 years prior to M\=/206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

In CA, the cultivars with broad adaptation and adoption over the
33-year period of this study are M\=/202, M\=/206, and M\=/105. Cold
tolerance is a major problem for much of the California rice producing
region ~\citep{espe2017}, although it is a bigger problem in the
southern portion of the Sacramento Valley due to the influence of the
ocean winds through the Delta ~\citep{espe16a}. These three cultivars
have a higher degree of cold tolerance, high yield potential, and
relatively short duration and are thus broadly grown with over 70\% of
the area in these cultivars at any given time during the study
period~\citep{ucce}. There are several cultivars that were found to
have better than average yields, even after correcting for yield
declines over time. However, these cultivars (M\=/201, M\=/204, M\=/205, and possibly
M\=/209; Fig. \ref{fig:ids_comp}) are not widely adopted. Importantly,
not all medium grain cultivars that were released were purported to
have higher yields than previous cultivars. Some cultivars had special
traits, but upon release had similar yields to more broadly adapted
cultivars. Examples of this include M\=/208, which had resistance to
blast disease (\textit{Magnaporthe grisea}) and M\=/402 which is a
premium quality cultivar. Other cultivars such as M\=/205 and M\=/209,
while high yielding, do not have the cold tolerance of the other
cultivars and are thus only recommended for the northern portion of
the Sacramento Valley.

@@

The high yielding variety M\=/202 was released in 1985 and became the
most widely planted cultivar for two decades~\citep{mckenzie2014}. At
the time of release, M\=/202 yields were reported at 11.0 Mg
ha\tsups{-1}~\citep{m202} which is almost identical to what we . This
value of 11.0 Mg ha\tsups{-1} is what we also found in this analysis
based on the average experimental yields prior to public release
(Fig. \ref{fig:first_yrs}). Since the release of M\=/202, the rice
yields of newly released varieties cultivars have remained between
10.8 and 11.5 Mg ha\tsups{-1} (Fig. \ref{fig:first_yrs}), with the
exception of M\=/204 and M\=/205 (M\=/204 is no longer grown and M\=/205 has
limited adaptation range so is only grown in a limited area). The
variety cultivar M\=/206 was released in 2004 and is currently still the
most widely grown rice cultivar in California. At the time of
its release, M\=/206 was reported to yield \Sexpr{round(9350*1.12/1000,1)} Mg
ha\tsups{-1} -- similar to what we found based on our analysis
(Fig. \ref{fig:first_yrs}). Based on this premise, it is clear that
improvements in yield potential have not been made over the 33-year
period of this study. In contrast, both~\citet{Omar2014} and
~\citet{mckenzie2014}, report yield potential gains in California rice
systems based on analyses that showed that newly released cultivars
had higher yields than previously released cultivars. Based on their
analyses,~\citet{Omar2014} reported yield gains from the rice breeding
program of 37 to 60 kg ha\tsups{-1} yr\tsups{-1} and
~\citet{mckenzie2014} a yield increase of 25 kg ha\tsups{-1}
yr\tsups{-1}. The simplest model (Model 2) in this study is similar
to the analyses of~\citet{Omar2014} and ~\citet{mckenzie2014}, and
results in similar findings; yields were estimated to increase by
\Sexpr{only_release_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:wo_erosion}), almost identical to estimates by
~\citet{mckenzie2014}.

The only way to reconcile the two apparently contradictory findings
above is that the yield of potential of cultivars decline over time
such that when newly released cultivars are compared with the older
cultivar (as shown by ~\citet{Omar2014}, ~\citet{mckenzie2014}, and
Fig. \ref{fig:wo_erosion}), yields of the new cultivar are
higher. Here we provide evidence that cultivar yield does decline - on
average at a rate of \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:RES_year}). When these declines are taken into account,
there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of a rate of yield increase equal to or greater than 25
kg ha\tsups{-1} yr\tsups{-1} during the study
period. ~\citet{Peng2010} also found that a cultivar's yield potential
declines over time and reported a 15\% decline over a 30 yr period.
In this study, over a 30-year period and a \Sexpr{yrs_coef*-1} kg
ha\tsups{-1} yr\tsups{-1} rate, the yield decline would be
approximately \Sexpr{round((yrs_coef*-30)/mean(mod_data$yield_kg),2)*100}\%.

@@

Yields of a particular cultivar may decline over time for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is not likely the case
here as the fields used in this study grow rice only one season a year
and tend to be non-flooded for the remainder of the year. Furthermore,
at this location and unlike most CA rice fields, the fields are left
fallow every other year. Secondly, small changes in biotic and abiotic
forces have previously been speculated as the cause for yield decline
over time of the cultivar IR8 at IRRI~\citep{Peng1998, Peng2000,
  Peng2010}. Since the cultivar's genetics remain stable from the time
of selection but the conditions the cultivar experiences over time are
continuously changing, accumulated changes too small to reliably
measure could plausibly lead to yield declines over
time~\citep{Peng1998, Peng2000, Peng2010, Mackay2011}. However, it is
not clear if this mechanism is responsible for the yield decline over
time of \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1} found in this
analysis. Biotic stressors such as pests and diseases are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure. Thirdly, changing climate may influence
yield potential with either warmer~\citep{van2013} or
cooler~\citep{espe2017} temperatures reducing
yields. Yield potential can be defined as the yield of optimum
cultivar in the absence of biotic and abiotic stresses in a given
environment \citep{vanIttersum2013}. However, ~\citet{espe2017} found
no substantial trends in growing season temperatures (either high,
low, or average temps) in the Sacramento Valley during the study
period, suggesting that older cultivars do not have lower yield
potential due to changing climatic factors. While our analysis offers
inconclusive evidence of differences between cultivars in rate of
yield decline (i.e., the CIs for the interactions overlapped with
each other and zero), visual examination suggests that cultivars may
be different in terms of rates of yield decline
(Fig. \ref{fig:all_cult}); however, we do not have sufficient data nor
the proper controls in the existing data set to fully test this
hypothesis. More data is needed to further explore the exact cause of
the yield declines observed here, as well to determine if different
cultivars have differential declines over time.

@@

\begin{comment}
\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg ha\tsups{-1}
yr\tsups{-1}. Yet the CI estimated from these data using the best
model suggests that there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg ha\tsups{-1} yr\tsups{-1}. There are several
explanations for this discrepancy in conclusions. First, \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. Second, these yield decreases over time
were not taken into account possibly because \citet{mckenzie2014}
concluded that there was no trend in the yield of the major cultivars
over time. However this conclusion is based on a flawed interpretation
of the results of their statistical test. Failing to reject the
hypothesis that the trend in yield equals zero does not equate to the
confirmation of no trend. Instead, failing to reject the hypothesis
could suggest the analysis lacks the statistical power to detect the
trend. Third, it
is not surprising that no trend in yield over time was detected given
that \citet{mckenzie2014} used data from several locations, yet
analyzed those data using a simple t-test. By using this simple
analysis, \citet{mckenzie2014} ignore systematic errors related to
differences between observations from different locations, including
different management systems, growing environments, and crop
timing. By contrast, the investigation presented here used a single,
well managed system and analyzed the data using a mixed effects model
which fully accounts for non-independence between similar
observations. In doing this, the analysis makes fewer assumptions and
more fully captures the structural relationships in the data. Hence,
the estimates presented here are likely more reliable.
\end{comment}

>>>>

To test the hypothesis, the influence of breeding and yield decline over time
were quantified on yield over time using three nested Bayesian hierarchical
models. First, yield was modeled as a response to the number of years since
the cultivar was officially released and the year of release using a
mixed-effects linear regression. To account for similarities between
years and between cultivars, year and cultivar were included in the
model as random effects (Model 1).

@@

Where $yield_i$ is the grain yield for cultivar $k$ in plot $i$ during
year $j$,
the Intercept is the average yield for the RES across cultivars and years, $Rate_{yield\
decline}$ is the rate of yield decline per
${years\ since\ release_{ki}}$ for cultivar $k$ in plot $i$
(mean-centered, i.e., negative for years prior to release, zero at the year of release),
$Rate_{yield\ improvement}$ is the rate of yield improvement per
year (${release\ year_{ik}}$), $\beta_{year_j}$ is the random effect of
year $j$, and $\beta_{cultivar_k}$ is the random effect of cultivar
$k$, and lastly $\beta_{cultivar_k:year_j}$ is the a random effect for
the interaction term
for cultivar $k$ in year $j$. Two variations of this
full model were used to further explore yield dynamics over
time. For comparison to estimates where yield decline over time is not taken
into account, the model was used as described above except omitting
the $Rate_{yield\ decline}$ term (Model 2). Lastly, a model was fit
to estimate if there were differences between cultivars in the rate of
yield decline over time by adding allowing the rate of decline to vary
by cultivar (i.e., adding
a $Rate_{yield\ decline_k} * years\ since\ release_{ki}$ term to the
model for the rate of yield decline for cultivar $k$) (Model 3).

@@

There was general agreement between the estimates from the Bayesian
and classical statistical approaches for all models, with the
estimates being similar between similar models. The Bayesian estimates
are presented here due to the inherent flexibility of Bayesian
inference, though the classical estimates can be found in the
supplemental material. Comparing the cultivars' yield during each's
initial time (average 3 years) in the experimental trials showed that across cultivars, yields ranged from
\Sexpr{round(plot_data["M207", "mean"], 2)} (M\=/207) to
\Sexpr{round(plot_data["M205", "mean"], 2)} Mg ha\tsups{-1} (M\=/205)
(Fig \ref{fig:first_yrs}). The relationship between the year
a cultivar was released and yield over the 33-year study period was
estimated to be slightly negative (\Sexpr{release_yr_coef}
kg ha\tsups{-1} yr\tsups{-1}) and does not provide
evidence of genetic yield improvements during this
period. Individual exceptions to this were two cultivars (M\=/204 and
M\=/205) which were observed to have higher yields than the others in
the raw data (Fig. \ref{fig:RES_year}).

@@

Based on Model 3, there was no evidence of consistent differences between cultivars in
the rate of yield decline over time (i.e., all cultivar specific rates
of decline overlapped with eachother and zero). Differences between cultivars in the decline
of yield over time, seen in the raw data (Fig. \ref{fig:all_cult}),
are therefore likely small and uncertain. Given this, the model excluding varying
rates of decline by cultivar (Model 2) is presented here. The yields
of each cultivar over time were evaluated relative to when the
cultivar was released using the full model which accounts for both
yield decline over time and yield improvements over time (Model 2). The average decrease
in yield over time following a cultivar's release
was estimated at \Sexpr{yrs_coef} kg ha\tsups{-1} year\tsups{-1} (90\%
CI: \Sexpr{RES_CI[2]} to \Sexpr{RES_CI[1]}) (Fig. \ref{fig:RES_year};
Table \ref{tbl:coef}). The oldest cultivar in the trials
(\Sexpr{gsub("M","M-",unique(mod_data[which.min(mod_data[,"release_yr"]),"id"]))})
is estimated to have lost an average of
\Sexpr{round(yrs_coef*-36/1000,1)} Mg ha\tsups{-1} since its
release. Likewise, an estimated 560 kg ha\tsups{-1} yield
advantage of M\=/206 over M\=/202 (the current and former predominant
cultivars, respectively), can be attributed solely to the decline
of M\=/202's yield over time (Figs. \ref{fig:first_yrs} \&
\ref{fig:compare_yrs}). Further supporting this estimate of yield
decline, both M\=/202 and M\=/205 yielded similarly at their respective year of
release (Fig. \ref{fig:compare_yrs}), despite the fact that M\=/202 was
released 19 years prior to M\=/206 (Fig. \ref{fig:compare_yrs}; Table
\ref{tbl:ids}).

@@

In CA, the cultivars with broad adaptation and adoption over the
33-year period of this study are M\=/202, M\=/206, and M\=/105. Cold
tolerance is a major problem for much of the California rice producing
region ~\citep{espe2017}, although it is a bigger problem in the
southern portion of the Sacramento Valley due to the influence of the
ocean winds through the San Francisco Bay ~\citep{espe16a} (see Fig. \ref{fig:map}
for the geographic orientation of the rice growing area relative to the
San Francisco Bay). These three cultivars
have a higher degree of cold tolerance, high yield potential, and
relatively short duration and are thus broadly grown with over 70\% of
the area in these cultivars at any given time during the study
period~\citep{ucce}. There are several cultivars that were found to
have better than average yields, even after correcting for yield
declines over time. However, these cultivars (M\=/201, M\=/204, M\=/205, and possibly
M\=/209; Fig. \ref{fig:ids_comp}) are not widely adopted. Importantly,
not all medium grain cultivars that were released were purported to
have higher yields than previous cultivars. Some cultivars had special
traits, but upon release had similar yields to more broadly adapted
cultivars. Examples of this include M\=/208, which had resistance to
blast disease (\textit{Magnaporthe grisea}) and M\=/402 which is a
premium quality cultivar. Other cultivars such as M\=/205 and M\=/209,
while high yielding, do not have the cold tolerance of the other
cultivars and are thus only recommended for the northern portion of
the Sacramento Valley.

@@

The high yielding variety M\=/202 was released in 1985 and became the
most widely planted cultivar for two decades~\citep{mckenzie2014}. At
the time of release, M\=/202 yields were reported at 11.0 Mg
ha\tsups{-1}~\citep{m202} which is almost identical to what we . This
value of 11.0 Mg ha\tsups{-1} is what we also found in this analysis
based on the average experimental yields prior to public release
(Fig. \ref{fig:first_yrs}). Since the release of M\=/202, the rice
yields of newly released varieties cultivars have remained between
10.8 and 11.5 Mg ha\tsups{-1} (Fig. \ref{fig:first_yrs}), with the
exception of M\=/204 and M\=/205 (M\=/204 is no longer grown and M\=/205 has
limited adaptation range so is only grown in a limited area). The
variety cultivar M\=/206 was released in 2004 and is currently still the
most widely grown rice cultivar in California. At the time of
its release, M\=/206 was reported to yield \Sexpr{round(9350*1.12/1000,1)} Mg
ha\tsups{-1} -- similar to what we found based on our analysis
(Fig. \ref{fig:first_yrs}). Based on this premise, it is clear that
improvements in yield potential have not been made over the 33-year
period of this study. In contrast, both~\citet{Omar2014} and
~\citet{mckenzie2014}, report yield potential gains in California rice
systems based on analyses that showed that newly released cultivars
had higher yields than previously released cultivars. Based on their
analyses,~\citet{Omar2014} reported yield gains from the rice breeding
program of 37 to 60 kg ha\tsups{-1} yr\tsups{-1} and
~\citet{mckenzie2014} a yield increase of 25 kg ha\tsups{-1}
yr\tsups{-1}. The simplest model (Model 2) in this study is similar
to the procedures used by~\citet{Omar2014} and ~\citet{mckenzie2014}, and
results in similar findings; yields were estimated to increase by
an average of \Sexpr{only_release_coef} kg ha\tsups{-1} yr\tsups{-1}
(Fig. \ref{fig:wo_erosion}), almost identical to estimates by
~\citet{mckenzie2014}.

The only way to reconcile the two apparently contradictory findings
above is that the yield of potential of cultivars decline over time
such that when newly released cultivars are compared with the older
cultivar (as shown by ~\citet{Omar2014}, ~\citet{mckenzie2014}, and
Fig. \ref{fig:wo_erosion}), yields of the new cultivar are greater
than those of the older cultivar. Here we provide evidence that
cultivar yield does decline - on average at a rate of \Sexpr{yrs_coef}
kg ha\tsups{-1} yr\tsups{-1} (Fig. \ref{fig:RES_year}). When these
declines are taken into account, it becomes clear that substantial
increases in yield over the study period are unlikely; there is only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of a rate of yield increase equal to or greater than 25 kg
ha\tsups{-1} yr\tsups{-1} during the study period. ~\citet{Peng2010}
also found that a cultivar's yield potential declines over time and
reported a 15\% decline over a 30 yr period.  In this study, over a
30-year period and a \Sexpr{yrs_coef*-1} kg ha\tsups{-1} yr\tsups{-1}
rate, the yield decline would be approximately
\Sexpr{round((yrs_coef*-30)/mean(mod_data$yield_kg),2)*100}\%.

@@

Yields of a particular cultivar may decline over time for a number of
reasons. First, ~\citet{Nie2009} reported yield declines in rice due
to continuously flooded conditions over a long time period which could
be overcome with improved N management. This is not likely the case
here as the fields used in this study grow rice only one season a year
and tend to be non-flooded for the remainder of the year. Furthermore,
at this location and unlike most CA rice fields, the fields are left
fallow every other year. Secondly, small changes in biotic and abiotic
forces have previously been speculated as the cause for yield decline
over time of the cultivar IR8 at IRRI~\citep{Peng1998, Peng2000,
  Peng2010}. Since the cultivar's genetics remain stable from the time
of selection but the conditions the cultivar experiences over time are
continuously changing, accumulated changes too small to reliably
measure could plausibly lead to yield declines over
time~\citep{Peng1998, Peng2000, Peng2010, Mackay2011}. However, it is
not clear if this mechanism is responsible for the yield decline over
time of \Sexpr{yrs_coef} kg ha\tsups{-1} yr\tsups{-1} found in this
analysis. Biotic stressors such as pests and diseases are low in this
Mediterranean climate. Furthermore, these fields are fallowed every
other year and the residues are burned, further lowering the potential
for pest and disease pressure. Thirdly, changing climate may influence
yield potential with either warmer~\citep{van2013} or
cooler~\citep{espe2017} temperatures reducing
yields, in effect decreasing the evironmental conditions required for
higher yield potential. Yield potential can be defined as the yield of optimum
cultivar in the absence of biotic and abiotic stresses in a given
environment \citep{vanIttersum2013}. However, ~\citet{espe2017} found
no evidence of sustained trends in growing season temperatures (either high,
low, or average temperatures) in the Sacramento Valley during the study
period, which does not support this theory that new cultivars have higher yield
potential in the face of declining environmental conditions. While our analysis offers
inconclusive evidence of differences between cultivars in rate of
yield decline (i.e., the CIs for the interactions overlapped with
each other and zero), visual examination suggests that cultivars may
be different in terms of rates of yield decline
(Fig. \ref{fig:all_cult}); however, we do not have sufficient data nor
the proper controls in the existing data set to fully test this
hypothesis. More data is needed to further explore the exact cause of
the yield declines observed here, as well to determine if different
cultivars have differential declines over time.

@@

\begin{comment}

\citet{mckenzie2014} explored the yield of rice in this same system
during a similar time period using similar data as these here, but
concluded that genetic rice yield potential has increased over this
time period at a rate of 25 kg ha\tsups{-1} yr\tsups{-1}. Yet the CI
estimated from these data using the best model suggests that there is
only a
\Sexpr{round(sum((raw_samples$release_yr_c*scl)>=25)/nrow(raw_samples)*100,1)}\%
probability of the rate of yield increase is equal to or greater than
25 kg ha\tsups{-1} yr\tsups{-1}. There are several explanations for
this discrepancy in conclusions. First, \citet{mckenzie2014} did not
account for yield decreases over time in their analysis. As shown
here, when these decreases are not taken into about, the estimated
increase in yield over time is estimated to be approximately the same
as that found by \citet{mckenzie2014}. Second, these yield decreases
over time were not taken into account possibly because
\citet{mckenzie2014} concluded that there was no trend in the yield of
the major cultivars over time. However this conclusion is based on a
flawed interpretation of the results of their statistical
test. Failing to reject the hypothesis that the trend in yield equals
zero does not equate to the confirmation of no trend. Instead, failing
to reject the hypothesis could suggest the analysis lacks the
statistical power to detect the trend. Third, it is not surprising
that no trend in yield over time was detected given that
\citet{mckenzie2014} used data from several locations, yet analyzed
those data using a simple t-test. By using this simple analysis,
\citet{mckenzie2014} ignore systematic errors related to differences
between observations from different locations, including different
management systems, growing environments, and crop timing. By
contrast, the investigation presented here used a single, well managed
system and analyzed the data using a mixed effects model which fully
accounts for non-independence between similar observations. In doing
this, the analysis makes fewer assumptions and more fully captures the
structural relationships in the data. Hence, the estimates presented
here are likely more reliable.
\end{comment}
